Cluster Difference Imaging Photometric Survey (CDIPS) II. TOI$\,$837: A Young Validated Planet in IC~2602

### 2020-07-22 paper/ms.tex
Update manuscript
RV limits on HEB scenarios

\begin{abstract}
  We report the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA TESS mission, the ESA Gaia mission, ground-based photometry
  from El Sauce and ASTEP400, and spectroscopy from CHIRON, FEROS,
  and Veloce.  We find that TOI$\,$837 is a $T=9.9$ mag
  G0/F9 dwarf in the southern open cluster IC 2602.  The star
  and planet are therefore $35^{+11}_{-5}$ million years old.
  Combining the transit photometry with a prior on the stellar
  parameters derived from the cluster color-magnitude diagram, we find
  that the planet has an orbital period of $8.3\,{\rm d}$ and is
  slightly smaller than Jupiter ($R_{\rm p} = 0.80^{+0.09}_{-0.03}
  \,R_{\rm Jup}$).  The transits are either grazing or nearly grazing.
  From radial velocity monitoring, we limit the minimum mass to less
  than 1.20 $M_{\rm Jup}$ (3$\sigma$).  Grazing transits are a cause
  for concern, as they are usually indicative of astrophysical false
  positive scenarios.  Our follow-up data show that such scenarios are
  highly unlikely.  Our combined multi-color photometry,
  high-resolution imaging, and radial velocities rule out hierarchical
  eclipsing binary scenarios.  Background eclipsing binary scenarios,
  though limited by speckle imaging, formally remain a 0.2\%
  possibility.  TOI 837.01 is therefore a validated adolescent
  exoplanet.  Further observations of its stellar obliquity,
  atmosphere, and mass should confirm its planetary nature beyond
  doubt, and could improve our understanding of how the physical and
  orbital properties of exoplanets change in time.
\end{abstract}

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=.49\textwidth]{f2b.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Ground-based follow-up photometry.} The data were acquired
    using the 0.36$\,$m at El Sauce and the 0.40$\,$m ASTEP400 at Dome
    C.  Black points are binned to 10-minute intervals.  The gray line
    is the best-fit model to the combined TESS and ground-based data.
    Red and blue lines show $2\sigma$ lower limits on the transit
    depths in the Cousins-R and Johnson-B bandpasses used to rule out
    specific false positive scenarios (see
    Section~\ref{subsec:colorphot}).
		\label{fig:groundphot}
	}
\end{figure}

@@

		\vspace{-0.3cm}
		\subfloat{
			\includegraphics[width=0.47\textwidth]{f3.pdf}
		}
	\end{center}
	\vspace{-0.6cm}
	\caption{
    {\bf Speckle-imaging and velocimetry of \tn.} {\it Top:} SOAR
    HRCam contrast limits derived from point-source injection-recovery
    experiments. Star A ($\Delta T=4.7$, $2.3''$ West) is detected,
    and is also a resolved Gaia source.  It is co-moving with \tn, and
    its parallax and on-sky position imply that it is physically
    separated from \tn\ by $6.6\pm 0.1\,{\rm pc}$.  {\it Middle:}
    Radial velocity (RV) measurements, with best-fit instrument
    offsets and jitter terms included.  The expected scatter from
    starspots based on $v\sin i$ and the photometric modulation
    amplitude is of order 300m$\,$s$^{-1}$.  {\it Bottom:} RV
    measurements phased to the orbital ephemeris of \pn.  The planet
    is not detected.  The black line shows a circular Keplerian orbit
    representing the $3\sigma$ upper mass limit.
		\label{fig:followup}
	}
\end{figure}

@@

\subsubsection{RVs}
The combined radial velocities from FEROS, CHIRON, and Veloce can be
used to detect massive bound companions orbiting \tn.  We searched for
the presence of planetary and stellar-mass companions using
\texttt{radvel} \citep{fulton_radvel_2018}.  We assumed circular
orbits, and performed two sets of fits.

% timmy.drivers.oneoff_drivers.get_mass_upper_limit
For the first fit, we assumed a restrictive prior on the period and
time of periapse, using the known ephemeris from the transit.  We then
fitted for the semi-amplitude, instrument offsets, and jitter
parameters.  This yielded a non-detection of the planet's orbit.  The
corresponding 3-$\sigma$ (99.7$^{\rm th}$ percentile) upper limit on
$M_{\rm p} \sin i$ is $1.20 M_{\rm Jup}$.  The data and corresponding
model are shown in Figure~\ref{fig:followup}.

The above exercise ruled out the possibility that the observed
eclipses are caused by a stellar-mass object orbiting \tn.  The lack
of a linear trend in the radial velocities, particularly in the FEROS
data, further constrains the presence of a hierarchical binary system.
The consistency of the Gaia DR2 RV ($17.44 \pm 0.64\,$\kms) with
the FEROS RVs ($18.0 \pm 0.1$\kms) in theory places an even stronger limit on 
linear trends, since the two observation epochs are separated by roughly
five years.
To derive limits for this second possibility, we placed a wide
log-uniform prior on the semi-amplitude and period ($\log K\ [{\rm
m\,s^{-1}}] \sim \mathcal{U}(1,10^5)$, $\log P\ [{\rm days}] \sim
\mathcal{U}(0.1, 10^{15})$).  We then fitted for the semi-amplitude,
period, time of conjunction, instrument offsets, and jitter
parameters.  We converted the resulting posterior in period and
semi-amplitude to minimum mass and semi-major axis assuming Kepler's
third law.  The resulting $3\sigma$ limits are shown in
Figure~\ref{fig:fpscenario}.
{\it LGB NOTE: the procedure in the preceding paragraph might be modified to
	instead convert the FEROS limit $|\dot{\gamma}| < 0.82\,{\rm m\,s}^{-1}\,{\rm
		day}^{-1}$ directly into a semi-amplitude limit.}

@@

We fitted for the observed depths in different bandpasses using a
similar machinery as described in Section~\ref{subsec:planet}.  We
fitted each ground-based transit individually for the planet-to-star
size ratio, the impact parameter, and a local quadratic trend (the
ephemeris was assumed from an initial fit of only the TESS data).  The
corresponding $2$-$\sigma$ lower limits on the transit depths in
Cousins-R and Johnson-B band light curves were 2.82 and 1.77$\,$ppt,
respectively, and are shown in Figure~\ref{fig:groundphot}.
Particularly in our Johnson-B light curve, the transit depth is
correlated with the mean and linear slope of the light curve: a
smaller depth is allowed if the data are fitted with a larger linear
slope and a larger mean.  Our quoted limits marginalize over these
correlations, and the depth measurement itself is nearly Gaussian.

@@

The reason the above statement is an insufficient argument against
BEBs is that \tn\ is not a randomly selected star---it was selected
because it shows eclipses.  Given a foreground star that shows
eclipses, the probability of a background star being present is much
greater than for an arbitrary foreground star.  The relevant
populations need to be modeled at the Monte Carlo level.  We opt to
use \texttt{VESPA} to perform this population modeling
\citep{morton_efficient_2012,vespa_2015}.

@@

This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
\tn\ was included on the TESS ``short-cadence'' target list in part
thanks to the Guest Investigator program of G.\ Sacco (G011265).
%

>>>>

\begin{abstract}
  We report the discovery of TOI 837.01 and its validation as a
  transiting planet.  We characterize the system using data from the
  NASA TESS mission, the ESA Gaia mission, ground-based photometry
  from El Sauce and ASTEP400, and spectroscopy from CHIRON, FEROS,
  and Veloce.  We find that TOI$\,$837 is a $T=9.9$ mag
  G0/F9 dwarf in the southern open cluster IC 2602.  The star
  and planet are therefore $35^{+11}_{-5}$ million years old.
  Combining the transit photometry with a prior on the stellar
  parameters derived from the cluster color-magnitude diagram, we find
  that the planet has an orbital period of $8.3\,{\rm d}$ and is
  slightly smaller than Jupiter ($R_{\rm p} = 0.80^{+0.09}_{-0.03}
  \,R_{\rm Jup}$).  The transits are either grazing or nearly grazing.
  From radial velocity monitoring, we limit the minimum mass to less
  than 1.20 $M_{\rm Jup}$ (3-$\sigma$).  Grazing transits are a cause
  for concern, as they are usually indicative of astrophysical false
  positive scenarios.  Our follow-up data show that such scenarios are
  highly unlikely.  Our combined multi-color photometry,
  high-resolution imaging, and radial velocities rule out hierarchical
  eclipsing binary scenarios.  Background eclipsing binary scenarios,
  though limited by speckle imaging, formally remain a 0.2\%
  possibility.  TOI 837.01 is therefore a validated adolescent
  exoplanet.  Further observations of its stellar obliquity,
  atmosphere, and mass should confirm its planetary nature beyond
  doubt, and could improve our understanding of how the physical and
  orbital properties of exoplanets change in time.
\end{abstract}

@@

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=.49\textwidth]{f2b.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Ground-based follow-up photometry.} The data were acquired
    using the 0.36$\,$m at El Sauce and the 0.40$\,$m ASTEP400 at Dome
    C.  Black points are binned to 10-minute intervals.  The gray line
    is the best-fit model to the combined TESS and ground-based data.
    Red and blue lines show 2-$\sigma$ lower limits on the transit
    depths in the Cousins-R and Johnson-B bandpasses used to rule out
    specific false positive scenarios (see
    Section~\ref{subsec:colorphot}).
		\label{fig:groundphot}
	}
\end{figure}

@@

		\vspace{-0.3cm}
		\subfloat{
			\includegraphics[width=0.47\textwidth]{f3.pdf}
		}
	\end{center}
	\vspace{-0.6cm}
	\caption{
    {\bf Speckle-imaging and velocimetry of \tn.} {\it Top:} SOAR
    HRCam contrast limits derived from point-source injection-recovery
    experiments. Star A ($\Delta T=4.7$, $2.3''$ West) is detected,
    and is also a resolved Gaia source.  It is co-moving with \tn, and
    its parallax and on-sky position imply that it is physically
    separated from \tn\ by $6.6\pm 0.1\,{\rm pc}$.  {\it Middle:}
    Radial velocity (RV) measurements, with best-fit instrument
    offsets and jitter terms included.  The expected scatter from
    starspots based on $v\sin i$ and the photometric modulation
    amplitude is of order 300m$\,$s$^{-1}$.  {\it Bottom:} RV
    measurements phased to the orbital ephemeris of \pn.  The planet
    is not detected.  The black line shows a circular Keplerian orbit
    representing the 3-$\sigma$ upper mass limit.
		\label{fig:followup}
	}
\end{figure}

@@

\subsubsection{RVs}
The radial velocities from FEROS, CHIRON, and Veloce can be
used to detect massive bound companions orbiting \tn.  We searched for
the presence of planetary and stellar-mass companions using
\texttt{radvel} \citep{fulton_radvel_2018}.  We assumed circular
orbits, and performed two sets of fits.

% timmy.drivers.oneoff_drivers.get_mass_upper_limit
For the first fit we set a prior on the period and
time of periapse using the known ephemeris from the transit.  We then
fitted for the semi-amplitude, instrument offsets, and jitter
parameters.  This yielded a non-detection of the planet's orbit.  The
corresponding 3-$\sigma$ (99.7$^{\rm th}$ percentile) upper limit on
$M_{\rm p} \sin i$ is $1.20 M_{\rm Jup}$.  The data and corresponding
model are shown in Figure~\ref{fig:followup}.

The above exercise ruled out the possibility that the observed
eclipses are caused by a stellar-mass object orbiting \tn.  The lack
of a linear radial velocity trend, particularly in the FEROS
data, further constrains the presence of a hierarchical binary system.
Fitting a line to the FEROS velocities, we found that any third body
in the system would need to induce a linear radial velocity trend with
amplitude $|\dot{\gamma}| < 0.82\,{\rm m\,s}^{-1}\,{\rm
day}^{-1}$, at 3-$\sigma$ confidence.  The baseline of the FEROS data
is 253 days; the agreement between the mean Gaia DR2 velocity ($17.44 \pm
0.64\,$\kms) and that from FEROS ($18.0 \pm 0.1$\kms) in theory places
an additional limit on linear trends, since the two observation
epochs are separated by roughly five years.

To place limits on the properties of a possible bound hierarchical companion, we fitted the radial
velocity data for a Keplerian orbit using wide priors on
the semi-amplitude and period: $\log K\ [{\rm m\,s^{-1}}] \sim
\mathcal{U}(1,10^5)$, and $\log P\ [{\rm days}] \sim \mathcal{U}(0.1,
10^{15})$.  We then fitted for the semi-amplitude, period, time of
conjunction, instrument offsets, and jitter parameters.  We converted
the resulting posterior in period and semi-amplitude to minimum mass
and semi-major axis assuming Kepler's third law.  The resulting
3-$\sigma$ limits are shown in Figure~\ref{fig:fpscenario}.

@@

We fitted for the observed depths in different bandpasses using a
similar machinery as described in Section~\ref{subsec:planet}.  We
fitted each ground-based transit individually for the planet-to-star
size ratio, the impact parameter, and a local quadratic trend (the
ephemeris was assumed from an initial fit of only the TESS data).  The
corresponding 2-$\sigma$ lower limits on the transit depths in
Cousins-R and Johnson-B band light curves were 2.82 and 1.77$\,$ppt,
respectively, and are shown in Figure~\ref{fig:groundphot}.
Particularly in our Johnson-B light curve, the transit depth is
correlated with the mean and linear slope of the light curve: a
smaller depth is allowed if the data are fitted with a larger linear
slope and a larger mean.  Our quoted limits marginalize over these
correlations, and the depth measurement itself is nearly Gaussian.

@@

The reason the above statement is an insufficient argument against
BEBs is that \tn\ is not a randomly selected star---it was selected
because it shows eclipses.  Given a foreground star that shows
eclipses, the probability of a background star being present is much
greater than for an arbitrary foreground star.  The relevant
populations need to be modeled at the Monte Carlo level.  We opt to
use \texttt{VESPA} to model the populations
\citep{morton_efficient_2012,vespa_2015}.

@@

This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
\tn\ was included on the TESS 2-minute target list in part
thanks to the Guest Investigator program of G.\ Sacco (G011265).
%
